---
title: "phenotype_data_analysis"
author: "rheaprag"
date: "2021-12-03"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

Histograms were to created to analyze the final merged phenotype data.

## Histogram with No Transformation

The original histogram created showed that the data was skewed to the right. 

```{}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(data.table))

#read in final clop phenotype file
clop<-fread("/home/rhea/account_wflow/phenos/clop/clop_CrCl_pheno_final.txt")
#assign CrCl column to variable
Creatinine_clop <- clop$CrCl
#create histogram
clop_hist <- hist(Creatinine_clop, main="Creatinine Clearance of Clopidogril", col="darkmagenta")

#read in final noac phenotype file 
noac<-fread("/home/rhea/account_wflow/phenos/noac/noac_CrCl_pheno_final.txt")
#rename column name to 'CrCl'
colnames(noac)[10] <- "CrCl"
#assign CrCl column to variable
Creatinine_noac <- noac$CrCl
#convert any MD's to NA
Creatinine_noac_num <- as.numeric(Creatinine_noac)
#create histogram
noac_hist <- hist(Creatinine_noac, main="Creatinine Clearance of NOAC drugs", col="gold")
```

```{r, echo = FALSE}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(data.table))

#read in final clop phenotype file
clop<-fread("/home/rhea/account_wflow/phenos/clop/clop_CrCl_pheno_final.txt")
#assign CrCl column to variable
Creatinine_clop <- clop$CrCl
#create histogram
clop_hist <- hist(Creatinine_clop, main="Creatinine Clearance of Clopidogril", col="darkmagenta")

#read in final noac phenotype file 
noac<-fread("/home/rhea/account_wflow/phenos/noac/noac_CrCl_pheno_final.txt")
#rename column name to 'CrCl'
colnames(noac)[10] <- "CrCl"
#assign CrCl column to variable
Creatinine_noac <- noac$CrCl
#convert any MD's to NA
Creatinine_noac_num <- as.numeric(Creatinine_noac)
#create histogram
noac_hist <- hist(Creatinine_noac_num, main="Creatinine Clearance of NOAC drugs", col="gold")
```

## Log Transformation

In order to fix this, I first applied log transformation to the phenotype files:
```{}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(data.table))

#read in final clop phenotype file
clop<-fread("/home/rhea/account_wflow/phenos/clop/clop_CrCl_pheno_final.txt")
#assign CrCl column to variable
Creatinine_clop <- clop$CrCl
#apply log transformation
Creatinine_c_log <- log(Creatinine_clop)
#create histogram
clop_hist <- hist(Creatinine_c_log, main="Creatinine Clearance of Clopidogril", col="darkmagenta")

#read in final noac phenotype file 
noac<-fread("/home/rhea/account_wflow/phenos/noac/noac_CrCl_pheno_final.txt")
#rename column name to 'CrCl'
colnames(noac)[10] <- "CrCl"
#assign CrCl column to variable
Creatinine_noac <- noac$CrCl
#convert any MD's to NA
Creatinine_noac_num <- as.numeric(Creatinine_noac)
#apply log transformation
Creatinine_noac_log <- log(Creatinine_noac_num)
#create histogram
noac_hist <- hist(Creatinine_noac_log, main="Creatinine Clearance of NOAC drugs", col="gold")
```

```{r, echo = FALSE}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(data.table))

#clop
clop<-fread("/home/rhea/account_wflow/phenos/clop/clop_CrCl_pheno_final.txt")
Creatinine_clop <- clop$CrCl
Creatinine_c_log <- log(Creatinine_clop)
clop_hist <- hist(Creatinine_c_log,
main="Creatinine Clearance of Clopidogril",
col="darkmagenta")

#noac 
noac<-fread("/home/rhea/account_wflow/phenos/noac/noac_CrCl_pheno_final.txt")
colnames(noac)[10] <- "CrCl"
Creatinine_noac <- noac$CrCl
Creatinine_noac_num <- as.numeric(Creatinine_noac)
Creatinine_noac_log <- log(Creatinine_noac_num)
noac_hist <- hist(Creatinine_noac_log,
main="Creatinine Clearance of NOAC drugs",
col="gold")
```

The log transformed data was skewed to the left, so rank normalization is applied instead.

## Rank Normalization

```{}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(data.table))

#rank normalize files
source("/home/rhea/account_wflow/rntransform.R")
source("/home/rhea/account_wflow/ztransform.R")

#read in final clop phenotype file
clop<-fread("/home/rhea/account_wflow/phenos/clop/clop_CrCl_pheno_final.txt")
#apply rn transformation
Creatinine_clop <- rntransform(clop$CrCl)
#create histogram
clop_hist <- hist(Creatinine_clop, main="Creatinine Clearance of Clopidogril", col="darkmagenta")

#read in final noac phenotype file
noac<-fread("/home/rhea/account_wflow/phenos/noac/noac_CrCl_pheno_final.txt")
#rename column to 'CrCl'
colnames(noac)[10] <- "CrCl"
#assign CrClcolumn to variable
Creatinine_noac <- noac$CrCl
#convert MD's to NA
Creatinine_noac_num <- as.numeric(Creatinine_noac)
#apply rn transform
Creatinine_noac_rn <- rntransform(Creatinine_noac_num)
#create histogram
noac_hist <- hist(Creatinine_noac_rn, main="Creatinine Clearance of NOAC drugs", col="gold")
```

```{r, echo = FALSE}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(data.table))

#rank normalize files
source("/home/rhea/account_wflow/rntransform.R")
source("/home/rhea/account_wflow/ztransform.R")

clop<-fread("/home/rhea/account_wflow/phenos/clop/clop_CrCl_pheno_final.txt")
Creatinine_clop <- rntransform(clop$CrCl)
clop_hist <- hist(Creatinine_clop, main="Creatinine Clearance of Clopidogril", col="darkmagenta")

noac<-fread("/home/rhea/account_wflow/phenos/noac/noac_CrCl_pheno_final.txt")
colnames(noac)[10] <- "CrCl"
Creatinine_noac <- noac$CrCl
Creatinine_noac_num <- as.numeric(Creatinine_noac)
Creatinine_noac_rn <- rntransform(Creatinine_noac_num)
noac_hist <- hist(Creatinine_noac_rn, main="Creatinine Clearance of NOAC drugs", col="gold")
```

These histograms showed a normal distribution. The phenotype data used for the remainder of the calculations will be rank-normalized.

Histograms are all saved under `/home/rhea/account_wflow/phenos/{drug}`.
